Environmental and geographic variables are effective surrogates for genetic variation in conservation planning

### 2016-10-06 code/rmarkdown/article.Rmd
+ bug fixes with incorporating separate pcas for each species

To describe the environmental characteristics of each planning unit, we obtained data for 19 bioclimatic variables across the extent of the study area [at $30^{\prime\prime}$ resolution; @r33]. We used this data set because it is commonly to map habitat suitability in conservation planning [@r493]. These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$). To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subjected to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1). The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers. We calculated the average value for each of the `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components in each planning unit, and used these values to characterise climatic variation among the planning units (Figure S2). Thus we constructed a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation. We expect planning units further apart from each other in this environmental space to contain individuals with increasingly different patterns in their adaptive loci.

>>>>

To describe the environmental characteristics of each planning unit, we obtained data for 19 bioclimatic variables across the extent of the study area [at $30^{\prime\prime}$ resolution; @r33]. We used this data set because it is commonly to map habitat suitability in conservation planning [@r493]. These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$). The average value in each planning unit for each of the 19 layers was calculated. To reduce dimensionality, for each species, the climatic values associated with the planning units occupied by the species were subject to principal components analysis (PCA; Table S1). On average, the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` of the variation. We then used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterise climatic variation among the planning units for each species (Figure S2). Thus we constructed a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space for each species--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation for that species. We expect planning units further apart from each other in this environmental space to contain individuals with increasingly different patterns in their adaptive loci.

### 2016-10-06 code/rmarkdown/supporting_information.Rmd
+ bug fixes with incorporating separate pcas for each species

```{r, fig.width=15, fig.height=3, fig.cap='Climatic variation. Each panel depicts variation based on a different principle component (PC). Squares represent planning units. The color of each planning unit denotes the average PC value of pixels inside it. Planning units with more similar colors have more similar climate regimes.', echo=FALSE}
do.call(
	grid.arrange,
		append(
		llply(grep('^env\\_.*$', names(grid.DF), value=TRUE), function(x) {
			grid.FPLY$plot_title <- paste0('PC ', substr(x, nchar(x), nchar(x)))
			ggplot() +
				geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
					fill='grey85', color='grey70') +
				geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
					group='group', fill=x), color='grey10') +
				theme_classic() +
				theme(
					axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
					panel.border=element_rect(color='black', fill=NA, size=1),
					strip.background = element_rect(fill='grey20'),
					strip.text = element_text(color='white', size=11)
				) +
				coord_cartesian(
					xlim=buffered.range(grid.FPLY$long, 0),
					ylim=buffered.range(grid.FPLY$lat, 0)
				) +
				scale_fill_gradientn(name='',
					colors=brewer.pal(9, 'RdYlBu'),
					guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
				) +
				facet_wrap(~ plot_title)
		}),
		list(ncol=4)
	)
)
```

@@

```{r, echo=FALSE, results='asis'}
## make results table showing Eigen values
curr.file <- tempfile(fileext='.tex')
pca.DF[['Eigen Value']] <- c(round(pca.DF[['Eigen Value']][-19],2), pvalString(pca.DF[['Eigen Value']][19], digits=2))
pca.DF[['Variation explained (%)']] <- c(round(pca.DF[['Variation explained (%)']][1:3],2), pvalString(pca.DF[['Variation explained (%)']][-1:-3], digits=2))
pca.DF[['Accumulative variation explained (%)']] <- as.character(round(pca.DF[['Accumulative variation explained (%)']],2))
x <- pca.DF %>%
	latex(file=curr.file, 
		rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, col.just=rep('c', ncol(pca.DF)), here=TRUE,
		colheads=c("\\parbox[t]{2.5cm}{\\centering \\bfseries{Principle Component}}", "\\parbox[t]{2cm}{\\centering \\bfseries{Eigen value}}",
			"\\parbox[t]{3cm}{\\centering \\bfseries{Variation explained (\\%)}}", "\\parbox[t]{4.5cm}{\\centering \\bfseries{Cumulative variation explained (\\%)}}"),
		caption=paste0('Summary of principle components analysis (PCA) on bioclimatic variation across the study area. The first ',as.character(as.english(surrogate.params.LST[[MODE]]$number.components)),' principle components (PCs) were used for subsequent analysis.')
	)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
```

>>>>

```{r, fig.width=15, fig.height=3, fig.cap='Climatic variation. Each panel depicts variation based on a principle component (PC) for a different species. Squares represent planning units. The color of each planning unit denotes the average PC value of pixels inside it. For a given species, the planning units with more similar colors have more similar climate regimes.', echo=FALSE}
plot(1)
```

@@

```{r, echo=FALSE, results='asis'}
## make table showing pca resuls
curr.file <- tempfile(fileext='.tex')
x <- spp.pca.DF %>%
	mutate(species=gsub('_', ' ', species, fixed=TRUE), percent_variation_explained=percent_variation_explained*100) %>%
	latex(
		file=curr.file,
		rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, col.just=rep('c', 2),
		colheads=c('\\bfseries{Species}', '\\bfseries{Variation explained (\\%)}'), here=TRUE,
		caption=paste0('Summary of principal component analyses fit to the climatic variation in the species\' ranges. Percentages correspond to the percent of variation described by the first ',as.character(as.english(surrogate.params.LST[[MODE]]$number.components)),' principle components.')
	)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='\\cline{1-', replacement='\\cline{3-4} \\cline{6-', fixed=TRUE) %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
```
